\chapter{Introduction} \label{Introduction}

The automotive industry is transitioning into the next phase of powertrain technology. As automotive manufacturers are forced to meet tightening fleet-wide emissions regulations, the electrified vehicle market share is increasing. The European Union (EU) has established ambitious policy \cite{EUL110/5} mandating that all new cars and vans sold in Europe be zero-emission by 2035 as part of a broader strategy to achieve climate neutrality by 2050. The EU aims to have electrified vehicles - both Battery Electric Vehicles (BEVs) and Plug-in Hybrid Electric Vehicles (PHEVs) - make up 80~\% of its automotive market share by 2030. In China, government subsidies and investment in battery technology are also yielding a rapid advancement in electrified vehicle adoption - primarily in the passenger car segment. Whilst the direction of the industry remains susceptible to geopolitical and market influences, it is clear that many future powertrains will rely on electrification.

To achieve high-efficiency and adhere to packaging constraints, modern electrified powertrains utilize high-speed and low load motors \cite{Cai2021}. These motors introduce new challenges regarding NVH (Noise, Vibration and Harshness) and the tribology of interacting conjunctions. The compact, lightweight and efficient powertrains operate under significantly different working conditions to conventional internal combustion engine (ICE) systems. They are subject to different underlying physics; such as regime of lubrication, dynamic response and magneto-mechanical interactions. This style of powertrain architecture therefore involves high-speed rolling element bearing operation in both the motor and transmission.

These bearings are crucial structural components and their dynamic response significantly affects the behaviour of the interconnected structures. The dynamic behaviour of the bearing governs the force transmission from the excitation source to the housing and structure. It also influences the tribological contact conditions between rollers and raceways.

With a trend towards cost saving zero-prototype development, the use of simulation tools in modern powertrain development is growing. Multi-system vehicle powertrain concepts are pushing complexity of simulation models and this requires accurate and robust component level understanding. Performance characteristics of the bearings, such as NVH, friction and wear must be accurately modelled at the development stage to ensure full system success.


\section{Research Aims and Objectives} \label{Research Questions}

To determine whether implicit elastohydrodynamic lubrication modelling is dynamically necessary for accurate high-speed rolling element bearing simulation in electrified powertrains, and to establish a computationally viable method for its implementation.

The overarching aim of this thesis is to investigate the interaction between tribology and dynamics in rolling element bearings, with particular focus on the significance of this multi-physics interaction in high-speed automotive powertrain applications. Particular attention is given to operating conditions where lubricant entrainment into the roller-race contact becomes substantial - up to 25~000~$rpm$ - such that conventional dry analyses may no longer be valid.

This work evaluates the influence of the elastohydrodynamic (EHL) film on bearing contact behaviour and system dynamics, and assesses the necessity of modelling this film implicitly within flexible multi‑body (FMBD) environments. In parallel, the research investigates the potential of artificial neural networks (ANNs) to deliver accurate and computationally efficient tribological predictions suitable for deployment in FMBD simulations.

\paragraph{Research Objectives}
\begin{enumerate}
	\item To quantify the influence of EHL film formation on contact load, deflection and stiffness of rolling element bearings at the high speed, low torque operating conditions representative of electrified vehicle powertrains. 
	\item To determine the extent to which implicit EHL film representation modifies system-level dynamic behaviour within am FMBD environment.
	\item To evaluate the accuracy and computational suitability of ANNs for predicting central EHL film thickness across a representative operating domain.
	\item To determine whether ANN-based surrogate models can provide a viable embedded alternative to conventional film thickness formulations within FMBD environments.
\end{enumerate}

\paragraph{Research Tasks}
\begin{enumerate}
	\item Design and manufacture a high-speed bearing test rig capable of operating up to the defined speed and load limits. Instrument the rig to measure bearing orbital motion and establish a repeatable data acquisition and signalling methodology. 
	\item Extract time-resolved kinematic boundary conditions from experimental measurements for use in tribological conjunction modelling.
	\item Develop a lubricated component-level bearing model, including implicit representation of the EHL film at the roller–race contact.
	\item Develop a coupled co-simulation framework, integrating the lubricated bearing model within an FMBD environment.
	\item Implement a numerical EHL model to compute central film thickness across the required operating domain.
	\item Generate a structured training dataset using regime-constrained sampling of the EHL data.
	\item Design, train and validate an ANN capable of computing central EHL film thickness for a wide range of the required input parameters.
	\item Embed the trained ANN within the FMBD model and benchmark computational efficiency and predictive accuracy against the numerical solution.
\end{enumerate}

\paragraph{Scope of the Research}
This research is confined to the investigation of lubricated rolling element bearings operating under high-speed, low-load conditions representative of electrified automotive powertrains. The study focuses specifically on cylindrical roller bearings in which lubricant entrainment at rotational speeds up to 25~000~$rpm$ significantly influences contact behaviour.

The work examines the influence of the EHL film on bearing load–deflection characteristics and the resulting modification of effective stiffness within a defined FMBD framework. The analysis is confined to smooth steel–steel contacts and the prediction of central film thickness as the governing tribological parameter affecting contact compliance. The system considered comprises a flexible shaft and rigid housing representative of a high-speed electric motor and single-stage gear arrangement.

Thermal effects, starvation, surface roughness interactions, and wear are beyond the scope of this research. The study does not seek to develop a comprehensive thermo-tribo-dynamic bearing model, but rather to determine whether implicit representation of EHL effects is necessary for accurate dynamic simulation.

ANNs are considered as surrogate models for central film thickness prediction, with emphasis on computational suitability for integration within dynamic simulations.

The scope is therefore limited to establishing the modelling boundaries required to capture the dynamic significance of EHL in high-speed rolling element bearings for electrified powertrain applications.

\section{Contributions to Knowledge} \label{Contribution to Knowledge Intro}

The main novelties and contributions to knowledge from this thesis are summarised below:

\begin{enumerate}
	\item A novel experimental test rig was designed and constructed to measure the kinematic motion of a bearing at rotational speeds and loads up to 15~000~$rpm$ and 750~$N$ respectively. The bearing orbital motion was measured and used for conjunction and component level tribological analysis. The methodology of coupling experimental test with numerical tribological models has not been previously reported in this manner at these speeds. The outcome of this demonstrated the requirement of implicitly modelling the EHL film in future high-speed applications, since it contributed to a 149\% contact load increase at 15~000~$rpm$ when compared to conventional dry analyses.
	
	\item A coupled co-simulation approach was established to implicitly consider the EHL film in roller bearings within a high-speed system-level FMBD model. The model, consisting of a flexible shaft and rigid housing, replicates the operating conditions of a 54~$kW$ permanent magnet synchronous motor (PMSM) coupled to a first stage gear pair, operating at speeds up to 21~000~$rpm$. This was the first time in open literature that an implicitly lubricated multi-physics bearing model has been considered in the context of electrified powertrain dynamics. The outcome of this demonstrated that the increased contact deflection, due to the lubricant film inclusion, increases total bearing stiffness by 24.9~\% at 21~000~rpm. This effectively behaves as a non-linear, speed dependant radial preload on the bearing. The contribution of this stiffness increased the natural frequency of the system, and hence affected NVH response. This proves the requirement to consider the EHL film implicitly for rolling element bearing modelling under the high speed, low load operating conditions of electrified powertrains.
	
	\item An ANN was trained to predict EHL central film thickness with input data calculated using a 1D numerical EHL model. A wide value range of input variables necessary for the EHL calculation were used to train the model. The value range was consistent with common machine element contacts, with its applicability also extending to gear pairs and cam contacts. The novel methodology of constraining the training data using the Greenwood regime ensured high data quality whilst only requiring 2~000 training data points. The ANN achieved an MSE of 7.65~$\times 10^{-5}~\mu \mathrm{m}^2$ when benchmarked against the numerical solution, whilst reducing calculation time by a factor of 1~570. This methodology will further contribute to the computational efficiency and accuracy of tribological ANNs.
	
	\item The ANN was embedded within an FMBD system-level model to calculate EHL film thickness and consider it implicitly in the evaluation of the bearing and system dynamics. The film thickness evaluation achieves the accuracy of the numerical model without the associated computational limitations. This modelling method of combining a component level tribological ANN within an FMBD system-level model has not been previously reported.
\end{enumerate}

This work was performed in collaboration with AVL List GmbH in order to disseminate the outcomes of this research in the development of commercial codes: AVL EXCITE\textsuperscript{TM} M and AVL EXCITE\textsuperscript{TM} Power Unit.

\section{Structure of the Thesis} \label{Structure of Thesis}

The research in this thesis is presented in the following structure:

\paragraph{Chapter 1} introduces the topic of the thesis and provides the specific research questions that the work aims to answer.

\paragraph{Chapter 2} provides a review of the literature pertinent to the thesis topic, as well as covering the necessary principles required to address the research questions.

\paragraph{Chapter 3} presents a high-speed experimental test rig which was used to obtain kinematic boundary conditions for tribological models. The importance of including the lubricant film in high-speed dynamic bearing models is assessed here. The governing equations for the thesis are also introduced in this chapter.

\paragraph{Chapter 4} integrates the tribological models used in Chapter 3 into a system-level FMBD model. Simulations up to 21~000~$rpm$ are performed using the dynamic input conditions of a 54~$kW$ PMSM coupled to a first stage gear pair. The EHL film is modelled implicitly at the roller-race conjunction, and its affect on system dynamics is evaluated.

\paragraph{Chapter 5} introduces an ANN  to address the shortcomings of the analytical approach when estimating EHL film thickness at high speeds. The ANN is trained using a wide design space of input variables. The accuracy of the model and computational efficiency is assessed. The ANN is embedded within the dynamic system level model from Chapter 4 as an alternative approach to the analytical solution for central film thickness. 

\paragraph{Chapter 6} presents the general conclusions of the work, and outlines potential future work that could follow this research.