% ************************** Thesis Abstract *****************************
% Use `abstract' as an option in the document class to print only the titlepage and the abstract.
%\begin{abstract}
\addcontentsline{toc}{chapter}{Publications}

\chapter*{Publications}

The results of this thesis are partially summarised in the following journal papers: %preprints:

\vspace{0.2cm}
\hspace{-1.7cm}(i) Questa, H., Mohammadpour, M., Theodossiades, S., Garner, C. P., Bewsher, S. R., \& Offner, G. (2021). Tribo-dynamic analysis of high-speed roller bearings for electrified vehicle powertrains. Tribology International, 154 (July 2020), 106675. \\https://doi.org/10.1016/j.triboint.2020.106675.

\vspace{0.2cm}
\hspace{-1.7cm}(ii) Questa, H., Mohammadpour, M., Theodossiades, S., Garner, C. P., Bewsher, S. R., \& Offner, G. (2023). Tribodynamic modelling of high-speed rolling element bearings in flexible multi-body environments. Machines, 11(1), 93. \\https://doi.org/10.3390/machines11010093.

\vspace{0.2cm}
\hspace{-1.7cm}(iii) Walker, J., Questa, H., Raman, A., Ahmed, M., Mohammadpour, M., Bewsher, S. R., \& Offner, G. (2023). Application of tribological artificial neural networks in machine elements. Tribology Letters, 71(1), 3. \\https://doi.org/10.1007/s11249-022-01673-5.

\vspace{1cm}
\hspace{-0.6cm}The results of this thesis are partially summarised in the following conference papers:

\vspace{0.2cm}
\hspace{-1.7cm}(i) Questa, H., Mohammadpour, M., Theodossiades, S., Garner, C. P., Bewsher, S. R., \& Offner, G. (2020). Tribodynamic modelling of high-speed rolling element bearings using experimentally obtained boundary conditions. Paper presented at NAFEMS 2020 UK Conference, United Kingdom.

\vspace{1cm}
\hspace{-0.6cm}Editorial features: %preprints:

\vspace{0.2cm}
\hspace{-1.7cm}(i) Tysoe, W. T., \& Spencer, N. D. (2023, February). Turning simulations into formulae by machine learning. Tribology and Lubrication Technology, 86–87. \\https://digitaleditions.walsworth.com/publication/?i=779793\&p=88\&view=issueViewer